Analysis of celiac disease

### 2018-03-15 5-results-celiac.Rmd
change cache

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.width = 10,
                      fig.asp = 0.7, dev = "svg", cache = TRUE)
options(width = 95)
```

@@

```{r}
# Compute PCA
obj.svd <- snp_autoSVD(G, CHR, POS, ncores = NCORES)
```

@@

```{r}
# Divide in training/test sets
dim(G)
set.seed(1)
ind.train <- sort(sample(nrow(G), size = 12e3))
ind.test <- setdiff(rows_along(G), ind.train)
```

@@

```{r}
# GWAS
system.time(
  gwas.train <- big_univLogReg(
    G, y[ind.train], ind.train = ind.train, 
    covar.train = obj.svd$u[ind.train, , drop = FALSE], 
    ncores = NCORES
  )
)
```

```{r, dev='png', fig.asp=0.8}
# Q-Q plot
snp_qq(gwas.train) +
  coord_cartesian(ylim = c(0, 15))
gwas.train.gc <- snp_gc(gwas.train)
```

```{r, dev="png", fig.asp=0.9, out.width="90%"}
# Manhattan plot
labels <- 1:22; labels[c(11, 13, 15, 17, 18, 20, 21)] <- ""
cowplot::plot_grid(
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels),
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels) +
    coord_cartesian(ylim = c(0, 22)) + 
    geom_hline(yintercept = -log10(5e-8), color = "red", linetype = 3),
  align = "hv", ncol = 1, labels = LETTERS[1:2], label_size = 25, scale = 0.95
)
```

```{r}
# Clumping on the test set
ind.keep <- snp_clumping(G, infos.chr = CHR,
                         ind.row = ind.test,
                         thr.r2 = 0.2, 
                         S = abs(gwas.train.gc$score),
                         size = 500,
                         is.size.in.bp = TRUE,
                         infos.pos = POS,
                         ncores = NCORES)
```

```{r}
# PRS
thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
lpS <- -predict(gwas.train.gc)
prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
               ind.test = ind.test,
               ind.keep = ind.keep,
               lpS.keep = lpS[ind.keep], 
               thr.list = thrs)
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
aucs <- apply(prs, 2, AUC, target = y[ind.test])
summary(aucs)
nb.pred[which.max(aucs)]
```

```{r}
pAUC <- function(pred, target, p = 0.1) {
  val.min <- min(target)
  q <- quantile(pred[target == val.min], probs = 1 - p)
  ind <- (target != val.min) | (pred > q)
  tryCatch(bigstatsr::AUC(pred[ind], target[ind]) * p, 
           error = function(e) p^2 / 2)  # all same prediction
}
plot_density_AUC <- function(pred, true = y[ind.test]) {
  data.frame(pred = pred, pheno = as.factor(true)) %>%
  myggplot() +
  geom_density(aes(pred, fill = pheno), alpha = 0.4) +
  ggtitle(sprintf("AUC = %s%% | pAUC = %s", 
                  round(100 * AUC(pred, true), 2),
                  round(pAUC(pred, true), 4))) + 
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
}
(p1 <- plot_density_AUC(prs[, which.max(aucs)]))
```

@@

```{r}
# Train logit-simple
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- apply(preds1, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds1.comb <- rowMeans(preds1)
AUC(preds1.comb, y[ind.test])
```

@@

```{r}
G2 <- big_attach("backingfiles/celiacQC_tripled1.rds")
system.time(
  logit2 <- big_spLogReg(X = G2, y01.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         ncores = NCORES)
)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", out.width = "70%", fig.width = 10,
                      fig.asp = 0.7, dev = "svg", cache = FALSE)
options(width = 95)
```

@@

```{r, cache=TRUE}
# Compute PCA
obj.svd <- snp_autoSVD(G, CHR, POS, ncores = NCORES)
```

@@

```{r, cache=TRUE}
# Divide in training/test sets
dim(G)
set.seed(1)
ind.train <- sort(sample(nrow(G), size = 12e3))
ind.test <- setdiff(rows_along(G), ind.train)
```

@@

```{r, cache=TRUE}
# GWAS
system.time(
  gwas.train <- big_univLogReg(
    G, y[ind.train], ind.train = ind.train, 
    covar.train = obj.svd$u[ind.train, , drop = FALSE], 
    ncores = NCORES
  )
)
```

```{r, dev='png', fig.asp=0.8, cache=TRUE}
# Q-Q plot
snp_qq(gwas.train) +
  coord_cartesian(ylim = c(0, 15))
gwas.train.gc <- snp_gc(gwas.train)
```

```{r, dev="png", fig.asp=0.9, out.width="90%", cache=TRUE}
# Manhattan plot
labels <- 1:22; labels[c(11, 13, 15, 17, 18, 20, 21)] <- ""
cowplot::plot_grid(
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels),
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels) +
    coord_cartesian(ylim = c(0, 22)) + 
    geom_hline(yintercept = -log10(5e-8), color = "red", linetype = 3),
  align = "hv", ncol = 1, labels = LETTERS[1:2], label_size = 25, scale = 0.95
)
```

```{r, cache=TRUE}
# Clumping on the test set
ind.keep <- snp_clumping(G, infos.chr = CHR,
                         ind.row = ind.test,
                         thr.r2 = 0.2, 
                         S = abs(gwas.train.gc$score),
                         size = 500,
                         is.size.in.bp = TRUE,
                         infos.pos = POS,
                         ncores = NCORES)
```

```{r, cache=TRUE}
# PRS
thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
lpS <- -predict(gwas.train.gc)
prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
               ind.test = ind.test,
               ind.keep = ind.keep,
               lpS.keep = lpS[ind.keep], 
               thr.list = thrs)
nb.pred <- sapply(thrs, function(thr) sum(lpS[ind.keep] > thr))
aucs <- apply(prs, 2, AUC, target = y[ind.test])
summary(aucs)
nb.pred[which.max(aucs)]
```

```{r}
pAUC <- function(pred, target, p = 0.1) {
  val.min <- min(target)
  q <- quantile(pred[target == val.min], probs = 1 - p)
  ind <- (target != val.min) | (pred > q)
  tryCatch(bigstatsr::AUC(pred[ind], target[ind]) * p, 
           error = function(e) p^2 / 2)  # all same prediction
}
plot_density_AUC <- function(pred, true = y[ind.test]) {
  true_fct <- forcats::fct_recode(as.factor(true), "Control" = "0", "Case" = "1")
  data.frame(Score = pred, Status = true_fct) %>%
  myggplot() +
  geom_density(aes(Score, fill = Status), alpha = 0.4) +
  ggtitle(sprintf("AUC = %s%% | pAUC = %s", 
                  round(100 * AUC(pred, true), 2),
                  round(pAUC(pred, true), 4))) + 
  theme(plot.title = element_text(size = rel(1.5), hjust = 0.5))
}
(p1 <- plot_density_AUC(prs[, which.max(aucs)]))
```

@@

```{r, cache=TRUE}
# Train logit-simple
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- apply(preds1, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds1.comb <- rowMeans(preds1)
AUC(preds1.comb, y[ind.test])
```

@@

```{r, cache=TRUE}
G2 <- big_attach("backingfiles/celiacQC_tripled1.rds")
system.time(
  logit2 <- big_spLogReg(X = G2, y01.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         ncores = NCORES)
)
